---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

if (requireNamespace("pkgload", quietly = TRUE)) {
  pkgload::load_all(".", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else {
  library(SelectBoost.FDA)
}
```

# SelectBoost.FDA

`SelectBoost.FDA` is an R package for variable selection in functional data
analysis. It combines FDA-native preprocessing and design objects with grouped
stability selection, interval summaries, FDA-aware `SelectBoost`, and a small
validation layer for simulation and benchmarking.

The package is designed for workflows where functional predictors are observed
on a grid, represented through basis expansions, or reduced to FPCA scores, and
where strong local or block-wise correlation makes ordinary variable selection
unstable.

## Main features

- FDA-native design objects built directly from raw curves, basis
  representations, FPCA scores, and scalar covariates.
- Train/test-safe preprocessing with identity transforms, standardization,
  spline-basis expansion, and FPCA.
- Grouped stability selection for functional blocks and interval summaries.
- FDA-aware `SelectBoost` wrappers plus a plain `SelectBoost` baseline.
- Simulation, benchmark, and evaluation helpers with mapped ground truth.

## Installation

You can install the development version from GitHub with:

```r
# install.packages("remotes")
remotes::install_github("bertran7/SelectBoost.FDA")
```

Some workflows rely on optional backends:

- `glmnet` for lasso-based grouped stability selection.
- `grpreg` for group lasso.
- `SGL` for sparse-group lasso.
- `FDboost` and `stabs` for the native `FDboost` stability-selection route.

## A first FDA-native workflow

The package ships with small example datasets so the full workflow can start
from raw functional inputs.

```{r}
data("spectra_example", package = "SelectBoost.FDA")

idx <- 1:30

design <- fda_design(
  response = spectra_example$response[idx],
  predictors = list(
    signal = fda_grid(
      spectra_example$predictors$signal[idx, ],
      argvals = spectra_example$grid,
      name = "signal",
      unit = "nm"
    ),
    nuisance = fda_grid(
      spectra_example$predictors$nuisance[idx, ],
      argvals = spectra_example$grid,
      name = "nuisance",
      unit = "nm"
    )
  ),
  scalar_covariates = spectra_example$scalar_covariates[idx, ],
  transforms = list(
    signal = fda_fpca(n_components = 3),
    nuisance = fda_bspline(df = 5)
  ),
  scalar_transform = fda_standardize(),
  family = "gaussian"
)

summary(design)
head(selection_map(design, level = "basis"))
```

## FDA-aware SelectBoost

`SelectBoost.FDA` extends `SelectBoost` with block-aware and region-aware
grouping while keeping the original perturbation engine.

```{r}
fit_sb <- fit_selectboost(
  design,
  mode = "fast",
  steps.seq = c(0.6, 0.3),
  c0lim = FALSE,
  B = 4
)

summary(fit_sb)
head(selection_map(fit_sb, level = "group", c0 = colnames(fit_sb$feature_selection)[1]))
```

## Grouped stability selection

Grouped stability selection is available through a common FDA interface. The
lasso route below requires the optional `glmnet` package.

```{r}
if (requireNamespace("glmnet", quietly = TRUE)) {
  fit_stab <- fit_stability(
    design,
    selector = "lasso",
    B = 8,
    cutoff = 0.5,
    seed = 1
  )

  summary(fit_stab)
  head(selection_map(fit_stab, level = "group"))
}
```

Interval summaries can be requested directly:

```{r}
if (requireNamespace("glmnet", quietly = TRUE)) {
  fit_interval <- interval_stability_selection(
    x = design,
    selector = "lasso",
    width = 4,
    B = 8,
    cutoff = 0.5,
    seed = 1
  )

  head(selection_map(fit_interval, level = "group"))
}
```

## Benchmarking on simulated FDA designs

The validation layer can be used to compare FDA-aware `SelectBoost` with a
plain `SelectBoost` baseline on the same simulated design and mapped truth.

```{r}
sim <- simulate_fda_scenario(
  n = 30,
  grid_length = 20,
  representation = "grid",
  seed = 1
)

bench <- benchmark_selection_methods(
  sim,
  methods = c("selectboost", "plain_selectboost"),
  levels = c("feature", "group"),
  selectboost_args = list(B = 3, steps.seq = 0.5, c0lim = FALSE),
  plain_selectboost_args = list(B = 3, steps.seq = 0.5, c0lim = FALSE)
)

head(bench$metrics)
```

## Further documentation

The package vignettes cover the main workflow families:

- discretized curves
- spectra and interval-aware `SelectBoost`
- basis and FPCA workflows
- methods, calibration, and formula interfaces
- simulation and benchmark workflows

## References

- Bertrand F., Aouadi I., Jung N., Carapito R., Vallat L., Bahram S., and
  Maumy-Bertrand M. SelectBoost: a general algorithm to enhance the performance
  of variable selection methods in correlated datasets. *Bioinformatics*.
  doi:10.1093/bioinformatics/btaa855
- Hofner B., Boccuto L., and Göker M. Stability selection and related
  subsampling-based selection procedures.
- Brockhaus S., Melcher M., Leisch F., and Greven S. FDboost:
  boosting functional regression models.
