---
title: "Basis and FPCA Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basis and FPCA Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
has_glmnet <- requireNamespace("glmnet", quietly = TRUE)
```

When functional predictors are already represented by spline coefficients or
FPCA scores, `SelectBoost.FDA` can still use the same design object and summary
machinery. The main difference is interpretive: scores/components are often the
units of interest rather than grid intervals.

## Build a design from FPCA scores

```{r}
library(SelectBoost.FDA)

set.seed(2026)
n <- 90

signal_scores <- cbind(
  rnorm(n),
  rnorm(n),
  rnorm(n)
)
noise_scores <- cbind(
  rnorm(n),
  rnorm(n)
)

y <- 2 * signal_scores[, 1] - 1.2 * signal_scores[, 2] + rnorm(n, sd = 0.35)

design <- fda_design(
  response = y,
  predictors = list(
    signal_fpca = fda_basis(
      signal_scores,
      basis_type = "fpca",
      component_names = paste0("PC", 1:3),
      name = "signal_fpca"
    ),
    noise_fpca = fda_basis(
      noise_scores,
      basis_type = "fpca",
      component_names = paste0("PC", 1:2),
      name = "noise_fpca"
    )
  ),
  family = "gaussian"
)

design
selection_map(design)
selection_map(design, level = "basis")
```

## Fit grouped stability selection

```{r, eval = has_glmnet}
fit <- fit_stability(
  design,
  selector = "glmnet",
  B = 30,
  sample_fraction = 0.5,
  cutoff = 0.6,
  seed = 7
)

fit
summary(fit)
selection_map(fit)
selection_map(fit, level = "basis")
plot(fit, type = "basis", value = "mean")
```

The basis-level summary is often the most convenient table for reporting:

- `n_components` counts the basis coefficients or FPCA scores in each predictor.
- `selected_components` reports how many exceed the stability threshold.
- `mean_feature_frequency` and `max_feature_frequency` summarize component-wise
  stability within each basis-expanded predictor.

## Interpretation note

For FPCA in particular, scores are often much less correlated than dense raw
grid points. In that setting, grouped stability selection is usually the more
natural default, while `SelectBoost` is often most useful for dense
discretizations or strongly correlated basis systems.
