---
title: "Discretized Curves and Grouped Stability Selection"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Discretized Curves and Grouped Stability Selection}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
has_grpreg <- requireNamespace("grpreg", quietly = TRUE)
```

`SelectBoost.FDA` can work directly with discretized functional predictors by
building an `fda_design` object and then fitting grouped stability selection on
top of it.

## Build a functional design

```{r}
library(SelectBoost.FDA)

set.seed(123)
n <- 80
grid <- seq(0, 1, length.out = 25)

signal_values <- outer(rnorm(n), sin(2 * pi * grid)) +
  matrix(rnorm(n * length(grid), sd = 0.2), nrow = n)
noise_values <- matrix(rnorm(n * length(grid)), nrow = n)

signal <- fda_grid(signal_values, argvals = grid, name = "signal", unit = "time")
noise <- fda_grid(noise_values, argvals = grid, name = "noise", unit = "time")
y <- 1.5 * signal_values[, 8] - signal_values[, 18] + rnorm(n, sd = 0.5)

design <- fda_design(
  response = y,
  predictors = list(signal = signal, noise = noise),
  family = "gaussian"
)

design
summary(design)
head(selection_map(design))
```

The design stores both the matrix passed to the fitting engine and the feature
map needed to come back to predictor names and domain locations later.

## Fit grouped stability selection

The following chunk is evaluated only when `grpreg` is installed.

```{r, eval = has_grpreg}
fit <- fit_stability(
  design,
  selector = "grpreg",
  B = 30,
  sample_fraction = 0.5,
  cutoff = 0.7,
  seed = 1
)

fit
summary(fit)
head(selection_map(fit))
plot(fit, type = "group")
fit$selected_groups
```

## Interval-level summaries

You can also move from predictor-level groups to non-overlapping intervals.

```{r, eval = has_grpreg}
interval_fit <- interval_stability_selection(
  design,
  width = 5,
  selector = "grpreg",
  B = 20,
  cutoff = 0.6,
  seed = 2
)

head(interval_fit$interval_table)
plot(
  interval_fit,
  type = "interval",
  value = "mean",
  facet = "predictor",
  legend_title = "Mean frequency",
  palette = grDevices::heat.colors(24)
)
interval_fit$selected_groups
```
