---
title: "Focused Benchmark Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Focused Benchmark Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The focused benchmark workflow is designed to demonstrate where FDA-aware
SelectBoost improves on plain SelectBoost for functional predictors: localized
dense signals, confounded blocks, high local correlation, and narrow active
regions.

The driver script writes only to an explicit `--output-dir` or, when omitted,
to `tempdir()`.

```{r, eval = FALSE}
system2(
  file.path(R.home("bin"), "Rscript"),
  c(
    "tools/run_focused_benchmark.R",
    "--quick",
    "--n-replicates=1",
    "--seed=20260616",
    paste0("--output-dir=", file.path(tempdir(), "selectboost_fda_focused_benchmark"))
  )
)
```

The package also exposes table extractors for report material.

```{r}
library(SelectBoost.FDA)

report_summary_table()
report_method_table()
report_formula_blocks()[c("selection_surface", "precision_recall")]
```

Saved sensitivity-study artifacts shipped with the package can be turned into a
compact benchmark table:

```{r}
report_benchmark_table(top_n = 5)
```

For new benchmark objects, use the renderer-neutral summary extractor:

```{r}
metrics <- data.frame(
  scenario = "localized_dense",
  representation = "grid",
  family = "gaussian",
  method = c("selectboost", "plain_selectboost"),
  level = "feature",
  precision = c(0.8, 0.6),
  recall = c(0.7, 0.6),
  f1 = c(0.746, 0.6),
  jaccard = c(0.59, 0.43),
  selection_rate = c(0.2, 0.3),
  stringsAsFactors = FALSE
)

as_benchmark_summary_data(metrics)
```
