---
title: "SelectBoost for Dense Spectra"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SelectBoost for Dense Spectra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
has_glmnet <- requireNamespace("glmnet", quietly = TRUE)
```

For dense spectra or finely sampled signals, `SelectBoost` is often most useful
when grouping is constrained by functional structure instead of using a purely
global correlation graph.

## Construct a spectral design object

```{r}
library(SelectBoost.FDA)

set.seed(321)
n <- 70
wavelength <- seq(1100, 2500, length.out = 40)

latent <- rnorm(n)
spectra_values <- outer(latent, exp(-((wavelength - 1700) / 180)^2)) +
  outer(rnorm(n), exp(-((wavelength - 2200) / 120)^2)) +
  matrix(rnorm(n * length(wavelength), sd = 0.15), nrow = n)

spectra <- fda_grid(
  spectra_values,
  argvals = wavelength,
  name = "spectra",
  unit = "nm"
)
y <- 2 * spectra_values[, 16] - 1.5 * spectra_values[, 30] + rnorm(n, sd = 0.4)

design <- fda_design(
  response = y,
  predictors = list(spectra = spectra),
  family = "gaussian"
)

head(selection_map(design))
```

## Fit FDA-aware SelectBoost

The fitting chunk below is evaluated only when `glmnet` is installed.

```{r, eval = has_glmnet}
sb_fit <- fit_selectboost(
  design,
  selector = "glmnet",
  selector_args = list(lambda_rule = "lambda.min"),
  mode = "fast",
  group_method = "threshold",
  bandwidth = 3,
  steps.seq = c(0.6, 0.2),
  B = 10,
  seed = 1
)

sb_fit
summary(sb_fit)
head(selection_map(sb_fit, c0 = colnames(sb_fit$feature_selection)[1]))
plot(
  sb_fit,
  type = "group",
  value = "mean",
  legend_title = "Mean selection",
  palette = grDevices::terrain.colors(24)
)
```

The `selection_map()` method keeps the wavelength information attached to each
coefficient, which is the main advantage of moving from raw matrices to
FDA-native design objects.
