---
title: "Simulation and Benchmark Workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation and Benchmark Workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
has_glmnet <- requireNamespace("glmnet", quietly = TRUE)
```

`SelectBoost.FDA` now includes a validation layer for repeated simulations,
method benchmarks, and plain-SelectBoost baselines.

## Simulate a benchmark scenario

```{r}
library(SelectBoost.FDA)

sim_grid <- simulate_fda_scenario(
  n = 60,
  grid_length = 30,
  representation = "grid",
  seed = 1
)

sim_grid
head(selection_map(sim_grid$design))
sim_grid$truth$active_predictors
```

The returned object keeps both the fitted `fda_design` and the mapped truth for
the transformed feature space.

## Benchmark multiple methods on shared truth

```{r, eval = has_glmnet}
bench <- benchmark_selection_methods(
  sim_grid,
  methods = c("stability", "interval", "selectboost", "plain_selectboost"),
  levels = c("feature", "group"),
  stability_args = list(selector = "lasso", B = 8, cutoff = 0.5, seed = 2),
  interval_args = list(selector = "lasso", width = 5, B = 8, cutoff = 0.5, seed = 3),
  selectboost_args = list(selector = "lasso", B = 4, steps.seq = c(0.7, 0.4), c0lim = FALSE),
  plain_selectboost_args = list(selector = "lasso", B = 4, steps.seq = c(0.7, 0.4), c0lim = FALSE)
)

bench
bench$metrics
head(selection_map(bench, level = "group"))
```

This keeps the comparison object available, so the same `selection_map()` and
`selected()` methods work on top of the benchmark output.

## Run a repeated study

```{r, eval = has_glmnet}
study <- run_simulation_study(
  n_rep = 2,
  simulate_args = list(
    n = 50,
    grid_length = 28,
    representation = "basis"
  ),
  benchmark_args = list(
    methods = c("stability", "selectboost", "plain_selectboost"),
    levels = c("feature", "group", "basis"),
    stability_args = list(selector = "lasso", B = 6, cutoff = 0.5, seed = 4),
    selectboost_args = list(selector = "lasso", B = 4, steps.seq = c(0.7, 0.4), c0lim = FALSE),
    plain_selectboost_args = list(selector = "lasso", B = 4, steps.seq = c(0.7, 0.4), c0lim = FALSE)
  ),
  seed = 10
)

study
summary(study)
```

The repeated-study summary reports the mean and standard deviation of recovery
metrics by method, evaluation level, and `c0` when applicable.
